The Part-Time Parliament\hfill\insertframenumber/\inserttotalframenumber

### 2011-05-24 disciplinas/usi_algsem/paxosTalk/paxosTalk.tex
23 pages paxosTalk

@@

\begin{frame}
  \frametitle{The Paxos algorithm}
  \framesubtitle{Some notations used}

@@

\begin{frame}
  \frametitle{The Paxos algorithm}
  \framesubtitle{Some more notations}

@@

\begin{frame}
  \frametitle{The Paxos algorithm}
  \framesubtitle{Still more notations\ldots}

@@

  \hspace{3 mm} \textbf{Phase 2b}) upon receiving \textit{Success}$(val)$, the process learns that the group achieved consensus on $val$.

@@

[1] Lamport, L. \textbf{The part-time parliament}, Technical Report 49, Systems Research Center, Digital Equipment Corp., 1989
\vspace{4 mm}

[2] Lamport, L. \textbf{Paxos Made Simple}, ACM SIGACT News, 2001
\vspace{4 mm}

[3] De Prisco, R., Lampson, B., Lynch, N. \textbf{Revisiting the Paxos Algorithm}, Distributed Algorithms, 1997
\vspace{4 mm}

[4] Lampson, B. \textbf{How to build a highly available system using consensus}, Distributed Algorithms, 1996
\vspace{4 mm}

[5] Castro, M., Liskov, B., \textbf{Practical Byzantine Fault Tolerance}, Operating Systems Design and Implementation, 1999

>>>>

%==========================================================
\section{About correctness}
%==========================================================

@@

\begin{frame}
%  \frametitle{The Paxos algorithm}
  \frametitle{Some notations used}

@@

\begin{frame}
%  \frametitle{The Paxos algorithm}
  \frametitle{Some more notations}

@@

\begin{frame}
%  \frametitle{The Paxos algorithm}
  \frametitle{Still more notations\ldots}

@@

  \hspace{3 mm} \textbf{Phase 3b}) upon receiving \textit{Success}$(val)$, the process learns that the group achieved consensus on $val$.

  \vspace{6mm}
  Every phase of this protocol is optional. Consistency is still maintained, although progress is not guaranteed.

  \vspace{6mm}
  Each process $q$ must keep \textit{MaxVote(b, q, $\beta$)} for each ballot numbered $b$ and remember all \textit{LastVote}$(b, v)$ messages it has sent.

%----------------------------------------------------------
\subsection{Basic protocol}
%----------------------------------------------------------

\begin{frame}
\frametitle{Basic protocol}

In the preliminary protocol, much information must be kept. A simpler, yet correct, protocol can be derived. Each process $p$ keeps then only:

\vspace{2 mm}
\mbox{\textit{\textbf{lastTried}}$[p]$ -- the number of the last ballot $p$ tried to initiate ($-\infty$ if none)}

\vspace{2mm}
\mbox{\textit{\textbf{prevVote}}$[p]$ -- the highest ballot for which $p$ ever voted ($-\infty$ if none)}

\vspace{2mm}
\textit{\textbf{nextBal}}$[p]$ -- the highest ballot number $b$ for which $p$ sent \textit{LastVote}$(b, v)$ ($-\infty$ if none)

\vspace{4mm}
In the preliminary protocol, \textit{LastVote}$(b, v)$ represented a ``promise'' from $q$ not to vote in any ballot numbered $b' : \text{\textit{LastVote}}(b, v)_{bal} < b' < b$

\vspace{2mm}
Here, \textit{LastVote}$(b, v)$ represents a stronger ``promise'' from $q$ not to vote in any ballot numbered $b' < b$
% although this promise is stronger, it just representing not sending a LastVote(...) in cases the preliminary prot. would. 
% as seen, this doesn't affect consistency
% ANYway, this is arguable, as one could see that the collection of promises from the preliminary are equivalent to this
% one single promise.

\begin{frame}
\frametitle{Basic protocol}

  Phase 1)

  \hspace{3 mm} \textbf{Phase 1a}) $p$ picks a ballot number $b > \text{\textit{lastTried}}[p]$ and sends \textit{NextBallot}$(b)$ to other processes

   %explain the if thing.
  \hspace{3 mm} \textbf{Phase 1b}) upon receipt of \textit{NextBallot}$(b)$ from $p$, \mbox{\textbf{if \textit{b}} > \textbf{\textit{nextBal}}[\textbf{\textit{q}}],} $q$ \mbox{sets $\text{\textit{nextBal}}[q]$ to $b$ and sends \textit{LastVote}$(b, v)$ to $p$, where $v = \text{\textit{prevVote}}[q]$}

  Phase 2)

  \hspace{3 mm} \textbf{Phase 2a}) after receiving \textit{LastVote}$(b, v)$ from a majority $Q$, \textbf{where \mbox{\textit{b $=$ lastTried}}}[\textbf{\textit{p}}], $p$ initiates ballot  number $b$ with value $val$ such that $B3(\beta)$ is maintained, by sending \textit{BeginBallot}$(b, val)$ to every process in $Q$
  %Bqrm = Q, Bval = val, Bbal = b

  \hspace{3 mm} \textbf{Phase 2b}) upon receiving \textit{BeginBallot}$(b, val)$ \textbf{with \textit{b $=$ nextBal}}[\textbf{\textit{q}}], $q$ votes on ballot number $b$, set $prevVote[q]$ to this vote and sends \textit{Voted}$(b, q)$ to $p$
  % here, each process first must promise to vote for the ballot B (and not receive any higher ballot B') to be able to vote for B...
  % in the preliminary, it didn't need any promise to respond in phase 2b

\begin{frame}
\frametitle{Basic protocol}

  Phase 3)

  \hspace{3 mm} \textbf{Phase 3b}) upon receiving \textit{Success}$(val)$, the process learns that the group achieved consensus on $val$.

  \vspace{6mm}
  Consistency is still maintained, and progress is still not guaranteed.

  \vspace{6mm}
  Anyway, it is impossible to guarantee termination of consensus in an asynchronous system [6]\ldots

%----------------------------------------------------------
\subsection{Complete (synchronous) protocol}
%----------------------------------------------------------

\begin{frame}
\frametitle{Complete (synchronous) protocol}

@@

[1] Lamport, L. \textbf{The part-time parliament}, Technical Report 49, Systems Research Center, Digital Equipment Corp., 1989

\vspace{4 mm}
[2] Lamport, L. \textbf{Paxos made simple}, ACM SIGACT News, 2001

\vspace{4 mm}
[3] De Prisco, R., Lampson, B., Lynch, N. \textbf{Revisiting the Paxos algorithm}, Distributed Algorithms, 1997

\vspace{4 mm}
[4] Lampson, B. \textbf{How to build a highly available system using consensus}, Distributed Algorithms, 1996

\vspace{4 mm}
[5] Castro, M., Liskov, B., \textbf{Practical Byzantine fault tolerance}, Operating Systems Design and Implementation, 1999

\vspace{4 mm}
[6] Fischer, M., Lynch, N., Paterson, M., \textbf{Impossibility of distributed consensus with one faulty process}, J. ACM 32 1 (Jan.), 374-382, 1985
